{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-28 thesis/tm.tex
more tm expts

\begin{enumerate}
	\item How quickly does Landslide find bugs in incorrect transactional programs? %of varying sizes?
	\item Does Landslide find any previously-unknown bugs in real-world transactional code?%
		\footnote{The savvy reader will realize that whether or not the author poses this evaluation question to begin with spoils its answer.}
	%\item How quickly does Landslide verify correct transactional programs of varying sizes?
	\item How does Iterative Deepening's (\sect{\ref{sec:quicksand-id}})
		performance compare to Maximal State Space mode (\sect{\ref{sec:landslide-quicksand-options}})?%
		\footnote{Not necessarily related to HTM,
		but the latter was implemented well after Chapter~\ref{chap:quicksand}'s conference paper was published,
		so this was the most convenient test suite to evaluate it on.}
	\item How well does Landslide's verification scale
		with increasing thread/iteration count for correct transactional programs?
	\item What further reduction can be achieved beyond that provided by
		the global-lock/failure-injection equivalences?
	%\item By the way, should MC research papers quantify variance in their CPU-time performance experiments?
\end{enumerate}

@@

\begin{itemize}
	\item Microbenchmarks
	\begin{itemize}
		\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}. %(assertion failure on {\tt count}).
		\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
		\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex locking failure path with an atomic {\tt xadd}, from \cite{tm-benchmark-cmu}.
		\item {\tt swap}: Microbenchmark that swaps values in an array, from \cite{tm-benchmark-cmu}.
		\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
	\end{itemize}
	\item Data structure tests
	\begin{itemize}
		\item {\tt avl\_insert}: AVL tree concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
		\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
		\item {\tt avl\_mutex}: {\tt avl\_fixed} with transactional sections simplified by abstraction into a mutex.
		\item {\tt map\_mutex}: {\tt map\_basic} simplified likewise.
	\end{itemize}
	\item Lock abstraction tests
	\begin{itemize}
		\item {\tt lock}(): Checks that multiple threads using a transactional lock
			cannot access the critical section simultaneously.
			% htm_spinlock and htm_mutex
		\item {\tt lock\_fast}(): Checks that a transactional lock's fast path will not suffer conflict aborts
			if its client threads' critical sections are independent.
			% htm_spinlock_fastpath and htm_mutex_fastpath
	\end{itemize}
		These are each parameterized over implementations {\tt spinlock} (from \cite{spinlock-rtm-github}),
		{\tt spin\_\allowbreak{}fixed} (spoilers!!),
		and {\tt mutex} (replaces the spinlock with a Landslide-annotated P2 mutex to reduce state space size).
\end{itemize}

@@

\begin{table}[h]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc|r|r}
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt htm2}
			& 2,1 & \cpu{16.30}	& \ints{22}	\\
			& 2,2 & \cpu{69.80}	& \ints{1970}	\\
			& 2,3 &\cpu{3510.00}	& \ints{104914}	\\
			%& 2,4 &	\ETAdag{...}	& \ETAdag{...}	\\
			& 3,1 &\cpu{41.07}	& \ints{941}	\\
			& 3,2 &	\ETAdag{9d 19h}	& \ETAdag{5058149}	\\ % add 10h ok
			% [JOB 4] progress: 923822/5058149 brs (18.264030%), ETA 9d 8h 30m 38s (elapsed 9h 59m 55s)
			& 4,1 & \cpu{12699.65}	& \ints{344240}	\\
			\hline
			{\tt counter}
			& 2,1 & \cpu{5.40}	& \ints{10}	\\
			& 2,2 & \cpu{10.91}	& \ints{190}	\\
			& 2,3 & \cpu{134.21}	& \ints{3970}	\\
			& 2,4 & \cpu{3009.40}	& \ints{86950}	\\
			& 3,1 & \cpu{8.30}	& \ints{120}	\\
			& 3,2 & \cpu{2038.90}	& \ints{60606}	\\
			& 4,1 & \cpu{101.95}	& \ints{3006}	\\
			\hline
			{\tt swap}
			& 2,1 & \cpu{38.79}	& \ints{99}	\\
			& 2,2 & \cpu{10299.27}	& \ints{233396}	\\
			& 3,1 & \ETAdag{todo} & \ETAdag{todo}	\\
			\hline
			{\tt avl\_insert}
			& 2,1 & \cpu{491.98}	& \ints{15056}	\\
			\hline
			{\tt avl\_fixed}
			& 2,1 & \cpu{572.49} 	& \ints{20420} 	\\
			& 2,2 & \ETAdag{todo}	& \ETAdag{todo} 	\\
			& 3,1 & \ETAdag{todo}	& \ETAdag{todo}	\\

@@

Note two curiosities:
firstly,
the impact that fixing Figure~\ref{fig:spinlockbug}'s performance bug
(changing a read+write to a read only)
had on even the correctness tests:
{\tt lock}({\tt spin\_fixed})'s state spaces were reduced by more than half compared to {\tt lock}({\tt spinlock}).
Secondly, {\tt spinlock} and {\tt spin\_fixed} take longer to test per interleaving
(CPU time divided by state space size)
than {\tt mutex},
because while {\tt mutex} abstracts away threads needing to wait their turn for the critical section
behind an API Landslide understands,
the spinlock's wait loop is open-coded, and Landslide must fall back on its costlier heuristic synchronization detection
(\sect{\ref{sec:landslide-blocking}}).
In this way (and also, of course, because {\tt mutex}'s state spaces are smaller overall),
{\tt mutex} can be seen in turn as a further abstraction reduction of {\tt spinlock}.

@@

			& 3,1 & \cpu{218.02} & \ints{185} \\ % certified causality ok
			& 3,2 & \ETAdag{todo} & \ETAdag{todo} \\
			& 4,1 & \\ % TODO

@@

			& 4,2 & \\ % TODO maybe?????? well its 10hrs

>>>>

\begin{enumerate}
	\item How quickly does Landslide find bugs in incorrect transactional programs? %of varying sizes?
	\item Does Landslide find any previously-unknown bugs in real-world transactional code?%
		\footnote{The savvy reader will realize that whether or not the author poses this evaluation question to begin with spoils its answer.}
	%\item How quickly does Landslide verify correct transactional programs of varying sizes?
	\item How does Iterative Deepening's (\sect{\ref{sec:quicksand-id}})
		performance compare to Maximal State Space mode (\sect{\ref{sec:landslide-quicksand-options}})?%
		\footnote{Not necessarily related to HTM,
		but the latter was implemented well after Chapter~\ref{chap:quicksand}'s conference paper was published,
		so this was the most convenient test suite to evaluate it on.}
	\item How well does Landslide's verification scale
		with increasing thread/iteration count for correct transactional programs?
	\item What further reduction can be achieved beyond the baseline provided by
		the global-lock/failure-injection equivalences?
	%\item By the way, should MC research papers quantify variance in their CPU-time performance experiments?
\end{enumerate}

@@

\begin{itemize}
	\item Microbenchmarks
	\begin{itemize}
		\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}. %(assertion failure on {\tt count}).
		\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
		\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex locking failure path with an atomic {\tt xadd}, from \cite{tm-benchmark-cmu}.
		\item {\tt swap}: Microbenchmark that swaps values in an array, from \cite{tm-benchmark-cmu}.
		\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
			% TODO: make sure figure number reference is right
		\item {\tt fig63}: Generalized version of Figure~\ref{fig:retry-set-example},
			contrived to induce as much reduction as possible from the retry sets optimization.
	\end{itemize}
	\item Data structure tests
	\begin{itemize}
		\item {\tt avl\_insert}: AVL tree concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
		\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
		\item {\tt avl\_mutex}: {\tt avl\_fixed} with transactional sections simplified by abstraction into a mutex.
		\item {\tt map\_mutex}: {\tt map\_basic} simplified likewise.
	\end{itemize}
	\item Lock abstraction tests
	\begin{itemize}
		\item {\tt lock}(): Checks that multiple threads using a transactional lock
			cannot access the critical section simultaneously.
			% htm_spinlock and htm_mutex
		\item {\tt lock\_fast}(): Checks that a transactional lock's fast path will not suffer conflict aborts
			if its client threads' critical sections are independent.
			% htm_spinlock_fastpath and htm_mutex_fastpath
	\end{itemize}
		These are each parameterized over implementations {\tt spinlock} (from \cite{spinlock-rtm-github}),
		{\tt spin\_\allowbreak{}fixed} (spoilers!!),
		and {\tt mutex} (replaces the spinlock with a Landslide-annotated P2 mutex to reduce state space size).
\end{itemize}

@@

\begin{table}[h]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc|r|r}
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt htm2}
			& 2,1 & \cpu{16.30}	& \ints{22}	\\
			& 2,2 & \cpu{69.80}	& \ints{1970}	\\
			& 2,3 &\cpu{3510.00}	& \ints{104914}	\\
			%& 2,4 &	\ETAdag{...}	& \ETAdag{...}	\\
			& 3,1 &\cpu{41.07}	& \ints{941}	\\
			& 3,2 &	\ETAdag{9d 19h}	& \ETAdag{5058149}	\\ % add 10h ok
			% [JOB 4] progress: 923822/5058149 brs (18.264030%), ETA 9d 8h 30m 38s (elapsed 9h 59m 55s)
			& 4,1 & \cpu{12699.65}	& \ints{344240}	\\
			\hline
			{\tt counter}
			& 2,1 & \cpu{5.40}	& \ints{10}	\\
			& 2,2 & \cpu{10.91}	& \ints{190}	\\
			& 2,3 & \cpu{134.21}	& \ints{3970}	\\
			& 2,4 & \cpu{3009.40}	& \ints{86950}	\\
			& 3,1 & \cpu{8.30}	& \ints{120}	\\
			& 3,2 & \cpu{2038.90}	& \ints{60606}	\\
			& 4,1 & \cpu{101.95}	& \ints{3006}	\\
			\hline
			{\tt swap}
			& 2,1 & \cpu{38.79}	& \ints{99}	\\
			& 2,2 & \cpu{10299.27}	& \ints{233396}	\\
			& 3,1 & \ETAdag{todo} & \ETAdag{todo}	\\
			\hline
			{\tt fig63}
			& 2,1 & \cpu{todo}	& \ints{todo}	\\
			& 2,2 & \cpu{todo}	& \ints{todo}	\\
			& 3,2 & \ETAdag{todo} & \ETAdag{todo}	\\
			\hline
			{\tt avl\_insert}
			& 2,1 & \cpu{491.98}	& \ints{15056}	\\
			\hline
			{\tt avl\_fixed}
			& 2,1 & \cpu{572.49} 	& \ints{20420} 	\\
			& 2,2 & \ETAdag{133y}	& \ETAdag{51484570} 	\\ % added 10hr but dosent matter
			% [JOB 19] progress: 408805/51484570 brs (0.794034%), ETA 132y 356d 8h 55m 16s (elapsed 9h 59m 19s)
			& 3,1 & \ETAdag{todo}	& \ETAdag{todo}	\\

@@

Note two curiosities:
firstly,
the impact that fixing Figure~\ref{fig:spinlockbug}'s performance bug
(changing a read+write to a read only)
had on even the correctness tests:
{\tt lock}({\tt spin\_fixed})'s state spaces were reduced by more than half compared to {\tt lock}({\tt spinlock}),
on account of DPOR no longer needing to reorder the (now) read-read access pairs.
Secondly, {\tt spinlock} and {\tt spin\_fixed} take longer to test per interleaving
%(CPU time divided by state space size)
than {\tt mutex}
(roughly 9 interleavings per second for the former, 28 for the latter),
because while {\tt mutex} abstracts away threads needing to wait their turn for the critical section
behind an API Landslide understands,
the spinlock's wait loop is open-coded, and Landslide must fall back on its costlier heuristic synchronization detection
(\sect{\ref{sec:landslide-blocking}}).
In this way (and also, of course, because {\tt mutex}'s state spaces are smaller overall),
{\tt mutex} can be seen in turn as a further abstraction reduction of {\tt spinlock}.

@@

			& 3,1 & \cpu{21.68} & \ints{186} \\ % certified causality ok
			& 3,2 & \ETAdag{8y 153d} & \ETAdag{1590416} \\ % caus ok
			% [JOB 4] progress: 187162/1590416 brs (11.768113%), ETA 8y 152d 16h 25m 38s (elapsed 9h 59m 49s)
			& 4,1 & \cpu{30408.58} & \ints {282084} \\ % causality ok

@@
